---
title: 'SpotAssayQuant Tutorial '
author: "Sarah Hawbaker"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("knitr")
library("EBImage")
library("kableExtra")
```

```{r, echo=FALSE}
df = data.frame(
  output = c("Spreadsheets", 
               "Outlines", 
               "Object Areas", 
               "Object Numbers", 
               "Rows", 
               "Graphs"), 
  module = c("ExportToSpreadsheet",
                          "SaveImages 1", 
                          "SaveImages 2", 
                          "SaveImages 3", 
                          "SaveImages 4", 
                          "N/A" ), 
  format = c(".csv", 
                    ".png", 
                    ".png", 
                    ".png", 
                    ".png", 
                    "N/A"))
```

## Overview {#overview}

This pipeline batch processes spot assay image sets and outputs all result data to .csv files.  In addition, it outputs a series of supplementary data and result images for spot checking.  You may find that most of this data is unecessary; it was made with a particularly problematic and inconsistent dataset in mind.  I have chosen to keep the auxillary data in for robust testing and verification.  See the [Input and Output](#io) section for more information. 

## Installation and Setup

### Installing Cellprofiler {#installation}

CellProfiler is available here:  https://cellprofiler.org/.  Click **Downloads >> CellProfiler**. Make sure to download the basic version and not CellProfiler Analyst.


```{r}
display(readImage("images/download1.png"))
```

### Choosing the Pipeline {#choosing}

Pipelines are located in the `Pipelines` folder.  There are two pipelines available, depending on the format of your input images.  For grayscale images, use the GrayScaleInput pipeline.  However, you may find that you want more control over the thresholding values and method.  In that case, there is a BinaryInput pipeline available.  Note that it is not difficult to modify the `Threshold` module in the GrayScaleInput pipeline, I have just created these two options for ease-of-use.  

For sample thresholding macros in ImageJ, see the `FIJI Macros` folder.


## Pipeline Setup

1.  **Open CellProfiler**.  Upon opening CellProfiler, you will see a black terminal window pop up and run a few lines of code.  This is just the .exe file launching.  Do not close this window while you are running the program, or it will exit.  I think it's best just to minimize it to avoid accidental termination.


```{r}
display(readImage("images/terminal.png"))
```

Close the introduction window.

2.  **Load the pipeline**.  Download your chosen pipeline from the `Pipelines` folder into a local directory.  Load into CellProfiler via **File >> Import >> Pipeline from File**. Your screen should look like this:

```{r}
display(readImage("images/overview.png"))
```


### Basic Navigation

The pipeline consists of a series of modules which each apply a different function to the image.  The module list is outlined in red below, and a single module in blue.  

```{r}
display(readImage("images/overview_highlights.png"))
```


```{r}
display(readImage("images/module settings.png"))
```



### Input and Output {#io}

Before we run the pipeline, we must make sure the output all goes to the correct place.  The output is currently structured as follows: 

* Spreadsheets: This is where all the raw data is stored.  

* Outlines:  This is a visual representation of the detected spots on each image. 

* Object Areas: A visual respresentation of measured data drawn on the spots themselves.  Currently this is set to "Area" but it can be changed in the `DisplayDataOnImage` module.  

* Object Numbers:  Similar to the Object Areas folder, but with the individual spot numbers displayed on each spot.  This is useful to verify that the rows are being grouped correctly. 

* Rows:  A visualization of the row groupings.  

* Graphs:  A place to store any user-generated graphs post-processing.  See the [Post Processing](#postprocessing) section for more information.


Should you not want this extra output, you can simply uncheck its associated module.  You do not need to worry about setting the output folder if the module is unchecked.

```{r, echo = FALSE}
kable(df, col.names = c("Output", "Associated Module", "File Format")) %>% 
  column_spec(1:3,border_left = TRUE, border_right = TRUE) %>% 
  kable_styling('bordered')

```


### Setting up the Output Folder:

1.  Click **View output settings**, which is located beneath the module list.  Set the **Default Output Folder** to your desired folder.  You do not need to change the **Output file format** setting.  If you want all your data to go into one folder, unsorted, you can skip the next step.  I don't recommend this as with large imagesets it will be messy to sort things out afterwards.

2.  For each of the associated modules, find the **Output File Location** setting.  Set to your desired output folder.  I like to create all empty folders ahead of time inside my main output folder.  There is an example output folder in the `Examples` folder for reference.

If you structure your output this way, select  **Default Output Folder sub-folder** from the drop-down menu and choose your folder from there.  Do this for each **SaveImages** module and the **ExportToSpreadsheet** module.


```{r, echo=FALSE}
display(readImage('images/output_subfolder.png'))
```



### Adding Input Files


Click on the **Images** module.  Drag and drop your files or folders into the pipeline.  To clear all files from the pipeline, go to **Edit >> Clear File List**.


### Defining the Grid

Before you run the pipline you will need to set up the "Grid" that determines how many rows and columns of spots you want to quantify.  Set this by clicking the **Define Grid** module and entering your values in the **Number of rows** and **Number of columns** fields.  The default settings are 7 and 4, respectively.

If you are sure that the image size and spot placements will be identical for each image, you can change the **Define a grid for which cycle?** setting to **Once**.  Otherwise it should stay at **Each Cycle**.


### Defining the Rows

CellProfiler allows you to create classifers that each spot can belong to.  For our purposes, we need the image to be grouped by rows.  Therefore each spot is assigned to a row depending on its y-coordinate value.  This allows us to calculate the total area for each row after the pipeline has finished running.

**Defining the Bins**

Click on the **Classify Objects** module.  Make sure all of your settings match the image below.  

```{r, echo = FALSE}
display(readImage('images/classify_settings.png'))
```


You will need change two fields.  First is the **Custom Threshold Values** field.  These are the values that define the bins.  Note that these values are for the **Center_Y** position, meaning that the center of the spot, not necessarily the whole spot, needs to fall inside the bin.  In the example below the values are set to `0,100,200,250,350,450,500` to define 7 rows.  

```{r,echo=FALSE}
display(readImage('images/classify_binned.png'))
```

Note that you do not need to set the upper bound, but you do need to include 0.

**Naming the Bins**

For however many rows you have, you will need that many names + 1.  I encourage you to stick to the naming convention shown in the example, where for 7 rows you have **Row1, Row2,...,Row7** as well a **Row0** that CellProfiler requires to put anything it is unable to classify.  These define the column headers for the .csv output file.

You will be able to see a visual representation of your row groupings in the ``Rows`` folder.  It's a good idea to spot check a few of these to make sure your row sums will be correct.

```{r, echo = FALSE}
display(readImage('images/example_rows.png'))
```




## Running the Pipeline


Click **Analyze Images**.  The pipeline will start. Typically it will take a moment or two to start up, but after that it will run relatively fast.  As it begins to process images, you will see windows pop up depending on which modules are set to visible.

### Define Grid

This is the first of two points in the pipeline where manual intervention is required. If you set the **Define a grid for which cycle?** setting to **Once** you will only need to do this step once.  Otherwise it will have to be done for each image.  

Click to set the coordinates for the top-left spot and the bottom-right spot.  Keep setting until you feel comfortable that the grid is nearly correct.  We can edit later, so it does not need to be perfect.

```{r, echo = FALSE}
display(readImage('images/grid_module.png'))
```

### Edit Objects

This is the second manual intervention. This module allows you to see the detected objects and edit them as needed.  Most images should not need any editing at all.  If you find that more than 10% need manual editing then either the settings are incorrect or this pipeline is not compatible with your dataset.  

The edit module has a "Show help" button which has detailed instructions for how to use it.  I will give an abridged version here, but I recommend to use the help module at first.  

Note that you cannot edit if the `pan` or `zoom` button is selected.  Always make sure to deselect after using.


The main commands you will need are: 

* Toggle.  Press **T** to toggle the display between full colors and outlines.  This will make the objects easier to see and edit.

* Edit mode:  **Right click** an object to go into edit mode. This will transform the object outline into a series of control points that you can then edit with the commands below.

* Delete:  Press **X** to delete a series of points by selection, **D** to delete a single point after clicking it. 

* Add:  Press **A** to add a control point at the location of your mouse after a single click.

* Join:  Press **J** to join two areas of control points.  Even if not contiguous the areas will be measured and counted as one object.

* Split:  Press **S** to split an object into two.  After pressing **S** you will click the two points that will form the dividing line.

```{r, echo=FALSE}
display(readImage('images/edit_module.png'))
```



## Post Processing {#postprocessing}

When the pipeline is completed you will see this pop-up:

```{r, echo=FALSE}
display(readImage('images/done.png'))
```

Go to the **Default Output Folder** to view your results.

You can use the [Spot Assay Sorting Tool](https://odonnelllab.shinyapps.io/CPGraphSortR/?_ga=2.42079597.937617316.1599767268-806153350.1599767268) to sort your output.  The source code is available [here](https://github.com/sah129/CPGraphSortR).  There is also a Jupyter Notebook version of the tool [here](https://github.com/sah129/Spot-Assay-Quant/blob/master/Scripts/generate_output.ipynb) that I have kept available for anyone interested.

```{r, echo = FALSE}
display(readImage('images/graphsort-1.png'))
```







